---
title: "Using the linreg Package"
author: "Christian Kammerer, Victor Guillo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the linreg Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

This vignette explains how to use the `linreg` function from the `packagelinreg` package. We will walk through using a simple example with the well-known `iris` dataset.

The `linreg` function implements a linear regression model that mimics the behavior of the base R `lm()` function, but adds additional plotting capabilities and summary statistics.

# Load necessary libraries

We begin by installing and loading the package.
```{r setup, echo=TRUE, message=FALSE, warning=FALSE, include=FALSE}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
```
```{r install-devtools, message=FALSE, warning=FALSE, results='hide'}
install.packages("devtools")
```
```{r load-libraries, echo=TRUE, message=FALSE, warning=FALSE}
devtools::install_github("Victorgl24/Advanced-Programming-R-Lab-4")
library(packagelinreg)

```
We load the data of our choice
```{r}
#Load dataset
data(iris)
head(iris)
```
And can now start fitting the model to the desired relation
```{r}
# Fitting the linear model using linreg package
model <- linreg$new(Petal.Length ~ Sepal.Length + Sepal.Width, data = iris)

```

Now that our model is fitted we have a wide range of functions that helps us
better understand our model
```{r}
# Display model summary
model$summary()

# Plot residuals vs fitted values
model$plot()

# Extract coefficients
coefs <- model$coef()
print(coefs)

# Get predicted values
predictions <- model$pred()
print(predictions)

#You can extract the coefficients of the model using the coef() function, which returns a named vector of coefficients. This is similar to the base lm() functionality.
coefs <- model$coef()
print(coefs)

predictions <- model$pred()
print(predictions)
```

A great video to better understand the concept of linear regression: https://www.youtube.com/watch?v=7ArmBVF2dCs <br> 
The documentation of the original lm package mentioned previously: https://www.rdocumentation.org/packages/stats/versions/3.6.2/topics/lm





